---
title: "Project 11 Analysis"
author: "Catherine C. Pollack"
date: "2/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache.lazy = FALSE)
```

# Bring in Libraries and Packages
```{r}
list_of_packages <- c("tidyverse", "janitor", "magrittr", "lmerTest", "lubridate", "ggsci", "corrplot", "ggpubr", "mediation", "reshape2", "lavaan")
new_packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) installed.packages(new_packages)

for (lib in list_of_packages) {
  library(lib, character.only = TRUE)
}
```

# Bring in Data
```{r}
dat <- read_csv("https://raw.githubusercontent.com/COVID19-DVRN/Project-11-Sick-Leave-Policy-in-the-US-and-Compliance-with-Social-Distancing-Measures/master/outputs/data/segmented_regression_files_with_google_mobility_by_date/01_03_02_all_dates_combined_economic_indicators_with_social_distancing_outcomes_state_wide.csv") %>%
  clean_names()
```

# Adding Variables
```{r}
dat %<>%
  mutate(date_of_state_emergency = current_date - day_since_state_emergency,
         date_of_statewide_neb_closure = current_date - day_since_statewide_neb_closure,
         date_of_statewide_stay_at_home = current_date - day_since_statewide_stay_at_home)

dat$weekday <- wday(dat$current_date, label = TRUE, abbr = FALSE)

dat$is_weekend <- ifelse(dat$weekday == "Saturday" | dat$weekday == "Sunday", "yes", "no")

dat %<>%
  mutate(is_after_ffl = case_when(
    current_date < "2020-03-18" ~ "Before",
    current_date >= "2020-03-18" ~ "After"
  ))

dat$is_after_ffl <- factor(dat$is_after_ffl, levels = c("Before", "After"))
```

# Redoing Matthew's Data: Figure
```{r}
plot_no_sick_leave <- dat %>%
  filter(weekday != "Saturday" & weekday != "Sunday" & paid_sick_leave == "No") %>%
  ggplot(aes(x = current_date, y = workplaces_percent_change_from_baseline, group = state_code)) +
  geom_rect(aes(xmin = min(date_of_state_emergency), 
                xmax = max(date_of_state_emergency),
                ymin = 0,
                ymax = max(workplaces_percent_change_from_baseline),
                fill = "State of Emergency Declared")) +
  geom_rect(aes(xmin = min(date_of_statewide_stay_at_home, na.rm = TRUE), 
                xmax = max(date_of_statewide_stay_at_home, na.rm = TRUE),
                ymin = -50,
                ymax = -25,
                fill = "Stay-at-Home Orders")) +
  geom_rect(aes(xmin = min(date_of_statewide_neb_closure, na.rm = TRUE), 
                xmax = max(date_of_statewide_neb_closure, na.rm = TRUE),
                ymin = -25,
                ymax = 0,
                fill = "Non-Essential Business Closure")) +
  scale_fill_manual(values = c("#90EE90", "#FFCCCB", "#ADD8E6")) + 
  geom_line(color = "grey") +
  ylim(-80, 20) + 
  geom_vline(xintercept = as.Date("03-18-2020", format = "%m-%d-%Y")) +
  theme_classic() +
  labs(x = "Date",
       y = "Workplace Travel (% Change from Baseline)",
       fill = "Policies",
       title = "No Paid Sick Leave")

plot_sick_leave <- dat %>%
  filter(weekday != "Saturday" & weekday != "Sunday" & paid_sick_leave == "Yes") %>%
  ggplot(aes(x = current_date, y = workplaces_percent_change_from_baseline, group = state_code)) +
  geom_rect(aes(xmin = min(date_of_state_emergency), 
                xmax = max(date_of_state_emergency),
                ymin = 0,
                ymax = max(workplaces_percent_change_from_baseline),
                fill = "State of Emergency Declared")) +
  geom_rect(aes(xmin = min(date_of_statewide_stay_at_home, na.rm = TRUE), 
                xmax = max(date_of_statewide_stay_at_home, na.rm = TRUE),
                ymin = -50,
                ymax = -25,
                fill = "Stay-at-Home Orders")) +
  geom_rect(aes(xmin = min(date_of_statewide_neb_closure, na.rm = TRUE), 
                xmax = max(date_of_statewide_neb_closure, na.rm = TRUE),
                ymin = -25,
                ymax = 0,
                fill = "Non-Essential Business Closure")) +
  scale_fill_manual(values = c("#90EE90", "#FFCCCB", "#ADD8E6")) + 
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("03-18-2020", format = "%m-%d-%Y")) +
  theme_classic() +
  ylim(-80, 20) + 
  labs(x = "Date",
       y = "Workplace Travel (% Change from Baseline)",
       fill = "Policies",
       title = "Paid Sick Leave")

ggarrange(plot_no_sick_leave, plot_sick_leave, 
          common.legend = TRUE,
          legend = "bottom")
ggsave("210416_mobility_policies_sickleave_noweekend.tiff", width = 7.25, height = 4.51)

```

# Second version of figure: same plot different color
```{r}
dat %>%
  filter(weekday != "Saturday" & weekday != "Sunday") %>%
  ggplot(aes(x = current_date,
             y = workplaces_percent_change_from_baseline, 
             color = paid_sick_leave, 
             group = state_code)) +
  geom_line() +
  scale_color_manual(values = c("lightgrey", "darkblue")) +
  geom_vline(xintercept = as.Date("03-18-2020", format = "%m-%d-%Y")) +
  theme_classic() +
  labs(x = "Date",
       y = "Workplace Travel (% Change from Baseline)",
       color = "State-Level Paid Sick Leave",
       title = "Mobility Changes Over Time by Paid Sick Leave")
ggsave("210416_mobility_policies_sickleave_noweekend_v2.tiff", width = 7.25, height = 4.51)
```

# Repeat for residential movement
```{r}
plot_no_sick_leave_residential <- dat %>%
  filter(weekday != "Saturday" & weekday != "Sunday" & paid_sick_leave == "No") %>%
  ggplot(aes(x = current_date, y = residential_percent_change_from_baseline, group = state_code)) +
  geom_rect(aes(xmin = min(date_of_state_emergency), 
                xmax = max(date_of_state_emergency),
                ymin = 0,
                ymax = max(residential_percent_change_from_baseline),
                fill = "State of Emergency Declared")) +
  geom_rect(aes(xmin = min(date_of_statewide_stay_at_home, na.rm = TRUE), 
                xmax = max(date_of_statewide_stay_at_home, na.rm = TRUE),
                ymin = -50,
                ymax = -25,
                fill = "Stay-at-Home Orders")) +
  geom_rect(aes(xmin = min(date_of_statewide_neb_closure, na.rm = TRUE), 
                xmax = max(date_of_statewide_neb_closure, na.rm = TRUE),
                ymin = -25,
                ymax = 0,
                fill = "Non-Essential Business Closure")) +
  scale_fill_manual(values = c("#90EE90", "#FFCCCB", "#ADD8E6")) + 
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("03-18-2020", format = "%m-%d-%Y")) +
  theme_classic() +
  labs(x = "Date",
       y = "Residential Travel (% Change from Baseline)",
       fill = "Policies",
       title = "No Paid Sick Leave")

plot_sick_leave_residential <- dat %>%
  filter(weekday != "Saturday" & weekday != "Sunday" & paid_sick_leave == "Yes") %>%
  ggplot(aes(x = current_date, y = residential_percent_change_from_baseline, group = state_code)) +
  geom_rect(aes(xmin = min(date_of_state_emergency), 
                xmax = max(date_of_state_emergency),
                ymin = 0,
                ymax = max(residential_percent_change_from_baseline),
                fill = "State of Emergency Declared")) +
  geom_rect(aes(xmin = min(date_of_statewide_stay_at_home, na.rm = TRUE), 
                xmax = max(date_of_statewide_stay_at_home, na.rm = TRUE),
                ymin = -50,
                ymax = -25,
                fill = "Stay-at-Home Orders")) +
  geom_rect(aes(xmin = min(date_of_statewide_neb_closure, na.rm = TRUE), 
                xmax = max(date_of_statewide_neb_closure, na.rm = TRUE),
                ymin = -25,
                ymax = 0,
                fill = "Non-Essential Business Closure")) +
  scale_fill_manual(values = c("#90EE90", "#FFCCCB", "#ADD8E6")) + 
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("03-18-2020", format = "%m-%d-%Y")) +
  theme_classic() +
  labs(x = "Date",
       y = "Residential Travel (% Change from Baseline)",
       fill = "Policies",
       title = "Paid Sick Leave")

ggarrange(plot_no_sick_leave_residential, plot_sick_leave_residential, 
          common.legend = TRUE,
          legend = "bottom")
ggsave("210310_residential_mobility_policies_sickleave_noweekend.tiff", width = 7.25, height = 4.51)

```

# Adding New Values
```{r}
dat$white <- dat$white * 100
dat$black <- dat$black * 100
dat$hispanic <- dat$hispanic * 100
dat$asian <- dat$asian * 100
dat$american_indian_alaska_native <- dat$american_indian_alaska_native * 100
dat$native_hawaiian_other_pacific_islander <- dat$native_hawaiian_other_pacific_islander * 100
dat$multiple_races <- dat$multiple_races * 100

dat$u_s_state_population_2019_estimate_numeric <- gsub("[\",]", "", dat$u_s_state_population_2019_estimate)
dat$u_s_state_population_2019_estimate_numeric <- as.numeric(dat$u_s_state_population_2019_estimate_numeric)

dat %<>%
  mutate(p_days_since_ffl = case_when(
    current_date >= "2020-03-18" ~ current_date - as.Date("2020-03-18"),
    TRUE ~ 0),
    population_adjusted_new_cases_per_100k = new_cases_since_a_week_before/as.numeric(u_s_state_population_2019_estimate_numeric) * 100000,
    population_adjusted_new_deaths_per_100k = new_deaths_since_a_week_before/as.numeric(u_s_state_population_2019_estimate_numeric) * 100000,
    population_adjusted_cumulative_cases_per_100k = cumulative_cases_at_current_date/as.numeric(u_s_state_population_2019_estimate_numeric) * 100000,
    
)

dat %<>%
  dplyr::select(state_code, current_date, unemployment_rate_feb_2020, unemployment_rate_mar_2020, unemployment_rate_apr_2020, unemployment_rate_may_2020, unemployment_rate_jun_2020, unemployment_rate_jul_2020) %>%
  melt(id = c("state_code", "current_date"), value.name = "unemployment_rate", variable.name = "unemployment_month") %>%
  mutate(unemployment_ismatch = case_when(
    month(current_date) == 2 & unemployment_month == "unemployment_rate_feb_2020" ~ 1,
    month(current_date) == 3 & unemployment_month == "unemployment_rate_mar_2020" ~ 1,
    month(current_date) == 4 & unemployment_month == "unemployment_rate_apr_2020" ~ 1,
    month(current_date) == 5 & unemployment_month == "unemployment_rate_may_2020" ~ 1,
    month(current_date) == 6 & unemployment_month == "unemployment_rate_jun_2020" ~ 1,
    month(current_date) == 7 & unemployment_month == "unemployment_rate_jul_2020" ~ 1,
    TRUE ~ 0
  )) %>%
  filter(unemployment_ismatch == 1) %>%
  right_join(., dat, by = c("state_code", "current_date"))

dat$month <- month(dat$current_date, label = TRUE)
```

## Correlation Matrix for Univariate Analysis -- Already checked and removed 
```{r}
univariate_analysis_continuous <- dat %>%
  select(cumulative_cases_at_current_date, cumulative_deaths_at_current_date, new_cases_since_a_week_before, new_deaths_since_a_week_before, retail_and_recreation_percent_change_from_baseline, grocery_and_pharmacy_percent_change_from_baseline, transit_stations_percent_change_from_baseline, residential_percent_change_from_baseline, raw_number_of_employees_in_thousands, unemployment_rate_feb_2020, unemployment_rate_jan_2021, median_income2017, avg_commute_time, avg_commute_time_public_transport, commuting_to_work_workers_16_years_and_over, commuting_to_work_workers_16_years_and_over_car_truck_or_van_drove_alone, percent_estimate_commuting_to_work_workers_16_years_and_over_car_truck_or_van_drove_alone, percent_estimate_commuting_to_work_workers_16_years_and_over_car_truck_or_van_carpooled, percent_estimate_commuting_to_work_workers_16_years_and_over_public_transportation_excluding_taxicab, percent_estimate_commuting_to_work_workers_16_years_and_over_walked, percent_estimate_commuting_to_work_workers_16_years_and_over_other_means, percent_estimate_commuting_to_work_workers_16_years_and_over_worked_at_home, labour_overall_index_score, wage_dimension_score, right_to_organize_dimension_score, mit_living_wage_2019, wage_ratio_living_wage_minimum_wage, worker_protection_dimension_2020_score, annual_state_gdp_for_2019, u_s_state_population_2019_estimate, u_s_state_land_area_square_miles, state_population_per_square_mile) 

corr <- cor(univariate_analysis_continuous)
corrplot(corr,
         diag = FALSE,
         tl.pos = "td", 
         tl.cex = 0.3,
         tl.col = "black",
         method = "color",
         type = "upper")
#write.csv(corr, file = "210216_correlations.csv")
```

## Univariate Analysis - Continuous
```{r}
univariate_analysis_continuous_list <- c("population_adjusted_cumulative_cases_per_100k", "population_adjusted_new_cases_per_100k", "population_adjusted_new_deaths_per_100k", "transit_stations_percent_change_from_baseline", "unemployment_rate", "median_income2017", "avg_commute_time", "avg_commute_time_public_transport", "percent_estimate_commuting_to_work_workers_16_years_and_over_car_truck_or_van_drove_alone", "percent_estimate_commuting_to_work_workers_16_years_and_over_car_truck_or_van_carpooled", "percent_estimate_commuting_to_work_workers_16_years_and_over_public_transportation_excluding_taxicab", "percent_estimate_commuting_to_work_workers_16_years_and_over_walked", "percent_estimate_commuting_to_work_workers_16_years_and_over_other_means", "percent_estimate_commuting_to_work_workers_16_years_and_over_worked_at_home", "labour_overall_index_score", "mit_living_wage_2019", "annual_state_gdp_for_2019", "u_s_state_land_area_square_miles", "state_population_per_square_mile", "black", "hispanic", "asian", "senior_population", "male_population", "poverty_rate")

for (var in univariate_analysis_continuous_list) {
  print(var)
  print(summary(model <- lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + get(var) + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no"))))
  print(confint(model))
  print("**********")
}
```

## Univariate analysis - Categorical
```{r}
dat[c(58:65)] <- lapply(dat[c(58:65)], function(x) {as.factor(as.character(x))})  ## as.factor() could also be used

dat %<>%
  mutate(election_results_coding_new = case_when(
    election_results_coding == "AlllR" ~ "AllR",
    TRUE ~ election_results_coding
  ))

dat$election_results_coding_new <- as.factor(dat$election_results_coding_new)
dat$election_results_coding_new <- relevel(dat$election_results_coding_new , ref = "Tie")
dat$dominant_sector_bls_feb_2020 <- relevel(as.factor(dat$dominant_sector_bls_feb_2020), ref = "Trade Transportation and Utilities")

for (var in colnames(dat)[c(20, 106, 58:65)]) {
  print(var)
  print(summary(model <- lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + get(var) + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no"))))
  print(confint(model))
  print("**********")
}
```

## Checking Correlations One More Time
```{r}
cor_matrix <- cor(dat[,univariate_analysis_continuous_list])
View(cor_matrix)
```


## Multivariable Categorical and Continuous - Paid Sick Leave, Bivariate Variable Selection
```{r}
summary(model <- lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + population_adjusted_new_cases_per_100k + population_adjusted_new_deaths_per_100k + unemployment_rate + median_income2017 + avg_commute_time + avg_commute_time_public_transport + labour_overall_index_score + mit_living_wage_2019 + annual_state_gdp_for_2019 + poverty_rate + dominant_sector_bls_feb_2020 + election_results_coding_new+ (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no")))

confint(model)
```

## Multivariable Categorical and Continuous - Paid Family Leave Leave, Bivariate Variable Selection
```{r}
summary(model <- lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_family_leave + population_adjusted_new_cases_per_100k + population_adjusted_new_deaths_per_100k + unemployment_rate + median_income2017 + avg_commute_time + avg_commute_time_public_transport + labour_overall_index_score + mit_living_wage_2019 + annual_state_gdp_for_2019 + poverty_rate + dominant_sector_bls_feb_2020 + election_results_coding_new+ (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no")))

confint(model)
```

## Multivariable Categorical and Continuous - Paid Sick Leave, Table Variables
```{r}
View(cor(dat %>% dplyr::select(population_adjusted_new_cases_per_100k, median_income2017, unemployment_rate, black, hispanic, asian, multiple_races , state_population_per_square_mile, senior_population, labour_overall_index_score)))
# Asian and multiple races correlated -- removing Multiple Races

summary(model <- lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + population_adjusted_new_cases_per_100k + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month), data = filter(dat, is_weekend == "no")))

confint(model)
```

## Multivariable Categorical and Continuous - Paid Family Leave, Table Variables
```{r}
summary(model <- lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_family_leave + population_adjusted_new_cases_per_100k + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month), data = filter(dat, is_weekend == "no")))

confint(model)
```

# Mediation Analysis
```{r}
fit_total_effect <- lmer(population_adjusted_new_cases_per_100k ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no"))
summary(fit_total_effect) #p= 0.30

#Need to add lme4:: to get mediation model to run
fit_mediator <- lme4::lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no"))
summary(fit_mediator) #p = 0.12
confint(fit_mediator)

#Need to add lme4:: to get mediation model to run
fit_dv <- lme4::lmer(population_adjusted_new_cases_per_100k ~ current_date + is_after_ffl + p_days_since_ffl + workplaces_percent_change_from_baseline + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no"))
summary(fit_dv)
confint(fit_dv)

mediation_final <- mediate(fit_mediator, 
                           fit_dv, 
                           treat = 'paid_sick_leave',
                           mediator = 'workplaces_percent_change_from_baseline')
summary(mediation_final)
```

# Residential Variable as a Secondary Outcome
```{r}
summary(model <- lmer(residential_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + population_adjusted_new_cases_per_100k + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no")))

confint(model <- lmer(residential_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + population_adjusted_new_cases_per_100k + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month),
                      data = filter(dat, is_weekend == "no")))

```

## Visualization
```{r}
dat_no_weekend <- dat %>%
  filter(is_weekend == "no")

dat_no_weekend$predicted_values_new_cases_contcat <- predict(lmer(workplaces_percent_change_from_baseline ~ current_date + is_after_ffl + p_days_since_ffl + paid_sick_leave + population_adjusted_new_cases_per_100k + median_income2017 + unemployment_rate + labour_overall_index_score + black + hispanic + asian + senior_population + state_population_per_square_mile + (1|state_code) + (1|month), data = filter(dat, is_weekend == "no")))

dat_no_weekend %>%
  mutate(paid_family_leave_plot = case_when(
    paid_family_leave == "No" ~ "No Paid Family Leave from State",
    paid_family_leave == "Yes" ~ "Paid Family Leave from State")) %>%
  ggplot(aes(x = current_date, y = predicted_values_new_cases_contcat, color = paid_family_leave_plot, group = state_code)) +
  geom_point(color = "grey") +
  geom_smooth(method = "lm") + 
  theme_classic() +
  facet_wrap(~paid_family_leave_plot) +
  scale_color_nejm() +
  labs(x = "Date",
       y = "Predicted New Daily Cases",
       color = "State-Level Paid Family Leave")
ggsave("210218_cases_policies_sickleave_predicted_contcat.tiff", width = 7.25, height = 4.51)
```
